Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
Typos

The numbers for uploading and downloading the different file vectors are
summarized in Tables \ref{tbl:hashingtimes:put1gb},
\ref{tbl:hashingtimes:get1gb}, \ref{tbl:hashingtimes:put100mb},
\ref{tbl:hashingtimes:get100mb}, \ref{tbl:hashingtimes:put1mb},
\ref{tbl:hashingtimes:get1mb}, \ref{tbl:hashingtimes:put1kb},
\ref{tbl:hashingtimes:get1kb}, \ref{tbl:hashingtimes:put1b} and
\ref{tbl:hashingtimes:get1b}.  The third column contains the time spent doing
other operations than hashing, i.e. \emph{Time spent not hashing = Total time -
Time spent hashing}. \emph{Fraction} displays the part of total time the calls
to the hashing libraries take.

@@

\paragraph{Modification to Tahoe-\ac{LAFS}.}
Because we write to two files after each call to {\verb digest }, an overhead in
the measurements that would not normally exist in Tahoe-\ac{LAFS}, will be
created. This overhead should be about the same for all candidates, since the
number of writes are equal. The results and calculations for this overhead,
based on runs with double \ac{SHA}-256, can be seen in Table
\ref{tbl:measure:overhead}.

@@

\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
mean implementations that does not conform to the \acp{KAT} supplied or
implementations which fail completely as hash functions, see
\autoref{sec:chosen_implementations} for details on the JH implementation.

@@

We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the {\verb -fPIC } flag.
This flag creates \emph{position independent code}, in other words machine
instruction code that can execute properly regardless of where in memory it
resides. This will make one less register available on the x86 architecture.
While we cannot directly prove that this will affect performance of the
implementations, we do however suspect it. A person writing an optimization for
\ac{SUPERCOP} will have to balance his use of resources, among them registers,
to achieve maximum speed.  The loss of one register will therefore give a
performance penalty.

@@

For large files, 1MB, 100MB and 1GB, we can see that the \ac{SHA}-3
implementations are outperforming double \ac{SHA}-256. Also for the large files
we can see that there is a performance penalty introduced by our bindings (Table
\ref{tbl:cython:overhead}), however the relative overhead are small
compared to the time spent in the underlying libraries. The reason for
this is that we are hashing larger sizes of data, especially for the 100MB and
1GB files (Tables \ref{tbl:tahoe:datauploaded} and
\ref{tbl:tahoe:datauploaded}). For these larger sizes the overhead caused by
using double \ac{SHA}-256 opposed to single is also relatively small and does
not cause a noteworthy performance penalty.

@@

If we look primarily at up- and download of the large files, we see that the candidates
that outperforms or are close to outperforming double \ac{SHA}-256, are BLAKE,
\ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
introduced by our bindings for the small files.

@@

\subsection{Time spent not hashing} A strange part of our result is that the
time Tahoe-\ac{LAFS} spends not callign the hash function is not static for all
candidates, which we expected.  The results of how much this deviance is
varies, although it looks to be bounded by some limits that are not the same
for all candidates. This leads us to believe that the hash function is to
blame, with noise from the system probably being the cause of the variation.
The time is recorded right before and right after the calls to our bindings,
which means that if the implementations are to be blamed they have to do
something to the system which delays later Tahoe-{LAFS} operations. We only
speculate in what they are changing and doing differently, but we suspect the
amount of registers and how much memory and cache the implementation uses as
probable causes.

@@

Even though few of the \ac{SHA}-3 implementations used by us managed to
outperform the implementation of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some
did for large files, which tells us there is a performance gain by switching
cryptographic hash function in Tahoe-\ac{LAFS}. \autoref{tbl:tahoe:gain}
compares the performance gains for BLAKE, which performed the best in our
measurements, versus \ac{SHA}-256, with the removal of overhead caused by our
measurements. We achieved as much as a 12.35\% gain when uploading the 100MB
files. The gain is smaller on downloads than uploads, this is
mostly caused by our Tahoe-\ac{LAFS} configuration, where we need to upload
shares to three servers, while we only need to download two of them to get the
original file.

@@

Four of the candidates have been shown to perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
Tahoe-\ac{LAFS} to upload a one gigabyte file, could improve by $5\%$ if the
best performing \ac{SHA}-3 candidate, BLAKE, were to be used. One has to bear in
mind that \ac{SHA}-256 has been optimized and tweaked for a number of years, and
that these performance differences undoubtedly will continue to increase as new
and better implementations will be developed for the final chosen \ac{SHA}-3
algorithm.

>>>>

The numbers for uploading and downloading the different test vectors are
summarized in Tables \ref{tbl:hashingtimes:put1gb},
\ref{tbl:hashingtimes:get1gb}, \ref{tbl:hashingtimes:put100mb},
\ref{tbl:hashingtimes:get100mb}, \ref{tbl:hashingtimes:put1mb},
\ref{tbl:hashingtimes:get1mb}, \ref{tbl:hashingtimes:put1kb},
\ref{tbl:hashingtimes:get1kb}, \ref{tbl:hashingtimes:put1b} and
\ref{tbl:hashingtimes:get1b}.  The third column contains the time spent doing
other operations than hashing, i.e. \emph{Time spent not hashing = Total time -
Time spent hashing}. \emph{Fraction} displays the part of total time the calls
to the hashing libraries take.

@@

\paragraph{Modification to Tahoe-\ac{LAFS}.}
Because we write to two files after each call to {\verb digest }, an overhead in
the measurements, that would not normally exist in Tahoe-\ac{LAFS}, will be
introduced. This overhead should be about the same for all candidates, since the
number of writes are equal. The results and calculations for this overhead,
based on runs with double \ac{SHA}-256, can be seen in Table
\ref{tbl:measure:overhead}.

@@

\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
mean implementations that does not conform to the \acp{KAT} supplied or
implementations which fail completely as hash functions, see
\autoref{sec:chosen_implementations} for details on the JH implementations.

@@

We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the {\verb -fPIC } flag.
This flag creates \emph{position independent code}, in other words machine
instruction code that can execute properly regardless of where in memory it
resides. This will make one less register available on the x86 architecture.
While we cannot directly prove that this will affect performance of the
implementations, we do however suspect it. A person writing an optimization for
\ac{SUPERCOP} will have to balance his use of resources, among them registers,
to achieve maximum speed.  The loss of one register should therefore give a
performance penalty.

@@

For large files, 1MB, 100MB and 1GB, we can see that the \ac{SHA}-3
implementations are outperforming double \ac{SHA}-256. Also for the large files
we can see that there is a performance penalty introduced by our bindings (Table
\ref{tbl:cython:overhead}), however the relative overhead are small
compared to the time spent in the underlying C implementation. The reason for
this is that we are hashing larger sizes of data, especially for the 100MB and
1GB files (Tables \ref{tbl:tahoe:datauploaded} and
\ref{tbl:tahoe:datadownloaded}). For these larger sizes the overhead caused by
using double \ac{SHA}-256 opposed to single is also relatively small and does
not cause a noteworthy performance penalty.

@@

If we look primarily at up- and download of large files, we see that the candidates
that outperforms or are close to outperforming double \ac{SHA}-256, are BLAKE,
\ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
introduced by our bindings for the small files.

@@

\subsection{Time spent not hashing}

A strange part of our result is that the
time Tahoe-\ac{LAFS} spends not calling the hash function is not static for all
candidates, which we expected.  The results of how much this deviance is
varies, although it looks to be bounded by some limits that are not the same
for all candidates. This leads us to believe that the hash function is to
blame, with noise from the system probably being the cause of the variation.
The time is recorded right before and right after the calls to our bindings,
which means that if the implementations are to be blamed they have to do
something to the system which delays later Tahoe-{LAFS} operations. We only
speculate in what they are changing and doing differently, but we suspect the
amount of registers and how much memory and cache the implementation uses as
probable causes.

@@

Even though few of the \ac{SHA}-3 implementations used by us managed to
outperform the implementation of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some
did for large files, which tells us there is a performance gain by switching
the cryptographic hash function in Tahoe-\ac{LAFS}. \autoref{tbl:tahoe:gain}
compares the performance gains for BLAKE, which performed the best in our
measurements, versus double \ac{SHA}-256, with the removal of overhead caused by our
measurements. We achieved as much as a 12.35\% gain when uploading the 100MB
files. The gain is smaller on downloads than uploads, this is
mostly caused by our Tahoe-\ac{LAFS} configuration, where we need to upload
shares to three servers, while we only need to download two of them to get the
original file.

@@

Four of the candidates have been shown to perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
Tahoe-\ac{LAFS} to upload a one gigabyte file, could improve by $12,36\%$ if the
best performing \ac{SHA}-3 candidate, BLAKE, were to be used. One has to bear in
mind that \ac{SHA}-256 has been optimized and tweaked for a number of years, and
that these performance differences undoubtedly will continue to increase as new
and better implementations will be developed for the final chosen \ac{SHA}-3
algorithm.
